Quantitative processing of broadband data as implemented in a scientific splitbeam echosounder

### 2022-02-18 Paper/Manuscript.tex

The power-budget equation (i.e., sonar equation) for a single target \citep[Formulation D, ][]{lunde2016} at frequency $\freqsym$ is:
%
\begin{equation}
\label{eq:TS}
\ts(\freqsym) = 10\log_{10}(\prxetf(\freqsym)) + 40\log_{10}(\range) + 2\absorp(\freqsym) \range 
- 10\log_{10}\left( \frac{\ptxe \wlen^2 \gain^2(\along_t,\athw_t,\freqsym)}{16\pi^2} \right),
\end{equation}
%

where $\prxetf(\freqsym)$ is the Fourier transform of the received electric power in a matched load for a signal from a single target at frequency $\freqsym$, $\range_t$ is the range to the target, $\absorp$ the acoustic absorption at frequency $f$, $\ptxe$ the transmitted electric power, $\wlen$ the acoustic wavelength, and $\gain$ the transducer gain along the main acoustic axis, incorporating beam compensation based on the estimated target bearing, $(\along_t,\athw_t)$.

@@

Lars: The peak in the send pulse needs to match with the peak in the signal.

Based on the point scattering strength samples and the phase angle samples, single targets can be detected, and range and bearing to the single targets can be estimated. This is typically achieved through a single echo detection algorithm (SED). Here we will assume that the samples from the pulse compressed data $\ypc(\samplesymt)$ originating from single target already have been identified, noting that the number of samples after the detected target may be higher than those those before the peak to include scattering processes for beyond ideal point targets. The alongship angle $\along_n$, athwartship angle $\athw_n$ and range $r_n$ at the \emph{peak} $\ypc(\samplesymt)$ (or peak $S_p(n)$?? current code uses $\ypc$) is used as estimates for $\along_t$, $\athw_t$ and $r_t$, respectively. A simple pseudo SED algorithm is implemented in the code for illustrative purposes.
%
\begin{figure}
\includegraphics[width=16cm]{Fig_singleTarget}
\caption{\label{fi:SED} The power (upper) and split beam angles (lower) for the single target. The vertical line corresponds to the range and angles for the single target.}
\end{figure}

@@

The discrete Fourier transforms of the target signal, $\ypctargetf(\samplesymf)$, and the reduced auto correlation signal, $\ymfautoredf(\samplesymf)$, are given by:
\begin{eqnarray}
\label{eq:DFT_Target_Auto}
\ypctargetf(\samplesymf) & = & \dft_\ndft(\ypctarget(\samplesymt)),\\
\ymfautoredf(\samplesymf) & = & \dft_\ndft(\ymfautored(\samplesymt)),
\end{eqnarray}
where $\dft$ indicates the Fourier transform of length $\ndft$ and $\samplesymf$ the sample index in the frequency domain.
The nomalized discrete Fourier transform of the target signal, $\ypctargetnormf(\samplesymf)$, (\ref{fi:SED}) is then calculated by: 
%
\begin{equation}
\label{eq:DFT_Target_Auto_Norm}
\ypctargetnormf(\samplesymf) = \frac{\ypctargetf(\samplesymf)} {\ymfautoredf(\samplesymf)}.
\end{equation}
%
The corresponding frequencies f\_m\_t (?) are calculate as follows... See the function EK80CalculationPaper.calcDFTforTS. Nils Olav did not quite get this so please enlighten me...
%
\begin{figure}
\includegraphics[width=16cm]{Fig_normalizedDFT}
\caption{\label{fi:SED}The $\ypctargetf(\samplesymf)$ (upper), $\ymfautoredf(\samplesymf)$ (middle) and $\ypctargetnormf(\samplesymf)$ (lower).}
\end{figure}

@@

Target strength can then be estimated using Eq. \ref{eq:TS}, but with $\freqsym$ replaced by the discrete index of frequency, $\samplesymf$:
\begin{equation}
\label{eq:TS_f}
\ts(\samplesymf) = 10\log_{10}(\prxetf(\samplesymf)) + 40\log_{10}(\range_t) + 2\absorp(\samplesymf)\range_t 
- 10\log_{10}\left( \frac{\ptxe \wlen_\samplesymf^2 \gain^2(\along_t,\athw_t,f_\samplesymf)}{16\pi^2} \right).
\end{equation}

Usikker på om vi skal bruka indeksar eller variablar for gain.

>>>>

The power-budget equation (i.e., sonar equation) for a single target \citep[Formulation D, ][]{lunde2016} at frequency $\freqsym$ is:
%
\begin{equation}
\label{eq:TS}
\ts(\freqsym) = 10\log_{10}(\prxetf(\freqsym)) + 40\log_{10}(\range) + 2\absorp(\freqsym) \range 
- 10\log_{10}\left( \frac{\ptxe \wlen^2 \gain^2(\along_t,\athw_t,\freqsym)}{16\pi^2} \right),
\end{equation}
%
where $\prxetf(\freqsym)$ is the Fourier transform of the received electric power in a matched load for a signal from a single target at frequency $\freqsym$, $\range_t$ is the range to the target, $\absorp$ the acoustic absorption at frequency $f$, $\ptxe$ the transmitted electric power, $\wlen$ the acoustic wavelength, and $\gain$ the transducer gain incorporating both the on axis gain $\gain_0(\freqsym)=\gain(0,0,\freqsym)$ and the beam pattern based on the estimated target bearing $(\along_t,\athw_t)$.

@@

Based on the point scattering strength samples and the phase angle samples, single targets can be detected, and range and bearing to the single targets can be estimated. This is typically achieved through a single echo detection algorithm (SED). Here we will assume that the samples from the pulse compressed data $\ypc(\samplesymt)$ originating from single target already have been identified, noting that the number of samples after the detected target may be higher than those those before the peak to include scattering processes for beyond ideal point targets. The alongship angle $\along_n$, athwartship angle $\athw_n$ and range $r_n$ at the \emph{peak} power \prxe(\samplesymt) within the detected target is used as estimates for $\along_t$, $\athw_t$ and $r_t$, respectively. A simple pseudo SED algorithm is implemented in the code for illustrative purposes.
%
\begin{figure}
\includegraphics[width=16cm]{Fig_singleTarget}
\caption{\label{fi:SED} The \prxe(\samplesymt) power (upper) and split beam angles ($\along_t$ and $\athw_t$) (lower) for the single target. The vertical line corresponds to the range $r_t$ for the single target.}
\end{figure}

@@

The discrete Fourier transforms of the target signal, $\ypctargetf(\samplesymf)$, and the reduced auto correlation signal, $\ymfautoredf(\samplesymf)$, are given by:
\begin{eqnarray}
\label{eq:DFT_Target_Auto}
\ypctargetf(\samplesymf) & = & \dft_\ndft(\ypctarget(\samplesymt)),\\
\ymfautoredf(\samplesymf) & = & \dft_\ndft(\ymfautored(\samplesymt)),
\end{eqnarray}
where $\dft$ indicates the Fourier transform of length $\ndft$ and $\samplesymf$ the sample index in the frequency domain.
The nomalized discrete Fourier transform of the target signal, $\ypctargetnormf(\samplesymf)$, (\ref{fi:SED}) is then calculated by: 
%
\begin{equation}
\label{eq:DFT_Target_Auto_Norm}
\ypctargetnormf(\samplesymf) = \frac{\ypctargetf(\samplesymf)} {\ymfautoredf(\samplesymf)}.
\end{equation}
%
The corresponding frequencies f\_m\_t (?) are calculate as follows... See the function EK80CalculationPaper.calcDFTforTS. Nils Olav did not quite get this so please enlighten me...

@@

Target strength can then be estimated using Eq. \ref{eq:TS}, but with $\freqsym$ replaced by the discrete index of frequency, $\samplesymf$:
\begin{equation}
\label{eq:TS_f}
\ts(\samplesymf) = 10\log_{10}(\prxetf(\samplesymf)) + 40\log_{10}(\range_t) + 2\absorp(\samplesymf)\range_t 
- 10\log_{10}\left( \frac{\ptxe \wlen_\samplesymf^2 \gain^2(\along_t,\athw_t,f_\samplesymf)}{16\pi^2} \right).
\end{equation}
%
\begin{figure}
\includegraphics[width=16cm]{Fig_TS}
\caption{\label{fi:SED}The $\ypctargetf(\samplesymf)$ (upper), $\ymfautoredf(\samplesymf)$ (middle) and $\ypctargetnormf(\samplesymf)$ (lower).}
\end{figure}
